Data Prep: IUCN and AquaMaps

### 2016-02-12 data_explore/data_explore.Rmd
updated plots with newly cleaned data

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           col_types = 'cdddd', progress = TRUE)

@@

### define representative species from each quadrant, for over-plotting
# spp_quads <- c('q1' = 'Kajikia albida', 
#                'q2' = 'Galaxea cryptoramosa', 
#                'q4' = 'Acanthurus nigroris', 
#                'q4a' = 'Metanephrops australiensis', 
#                'q3' = 'Conus magnificus')

# spp_rando <- c("Lagocephalus lagocephalus",
#                "Makaira nigricans",
#                "Pentacheles laevis",
#                "Stereomastis suhmi",
#                "Stereomastis surda") 
# 
### truncate the species names for the representative species
# spp_list_label <- spp_list %>%
#   filter(sciname %in% spp_rando) %>%
#   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
#   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
#   dplyr::select(-g, -s)

@@

spp_list <- spp_list %>% mutate(data_poor = ifelse(occurcells < 10, '< 10', '>= 10'))
### mongo plot time
scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
                                         transform(popn_category = reorder(popn_category, category_score)),
                                       aes(x = area_ratio, y = sm_perc, 
                                           color = data_poor,
                                           size = lg_area_pct)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
  geom_point(alpha = .6)

### Manage scales for color and size 
scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
  scale_colour_manual(values = c('>= 10' = 'green4', 
                                 '< 10'  = 'orangered2')) + 
  scale_size_continuous(limits = c(0, 100),
                        breaks = c( 0,    25,    50,    75,    100), 
                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
  scale_x_continuous(expand = c(0, 0), limits = c(0, 100)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))

scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
  labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
#       title = 'Distribution alignment vs Extent alignment',
       color = 'AM model cells', size = '% ocean range')

@@

spp_list_quads <- spp_list1 %>%
  mutate(quad = NA,
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q1$sciname, 'q1', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q2$sciname, 'q2', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q3$sciname, 'q3', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
write.csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'), row.names = FALSE)

@@

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'count', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, 
            #size = 2, 
            color = 'grey20') +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Relative proportion of species per quadrant', 
#       title = 'Taxonomic group by quadrant',
       fill = 'Alignment')

@@

write.csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'), row.names = FALSE)

@@

### Summarize quadrant species list by species group (e.g. CORALS1)
spp_risk_q1 <- spp_list_q1 %>%
  group_by(popn_category) %>%
  summarize(n_spp = first(n_spp), n_spp_q1 = n())
spp_risk_q2 <- spp_list_q2 %>%
  group_by(popn_category) %>%
  summarize(n_spp = first(n_spp), n_spp_q2 = n())
spp_risk_q3 <- spp_list_q3 %>%
  group_by(popn_category) %>%
  summarize(n_spp = first(n_spp), n_spp_q3 = n())
spp_risk_q4 <- spp_list_q4 %>%
  group_by(popn_category) %>%
  summarize(n_spp = first(n_spp), n_spp_q4 = n())

### Join species group by quadrant into a single dataframe
spp_risk_quadrants <- spp_risk_q1 %>%
  full_join(spp_risk_q2, by = c('popn_category', 'n_spp')) %>%
  full_join(spp_risk_q3, by = c('popn_category', 'n_spp')) %>%
  full_join(spp_risk_q4, by = c('popn_category', 'n_spp')) %>%
  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
  mutate(quad = str_replace(quad, 'n_spp_', ''),
         pct_quad = n_quad/n_spp,
         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))

spp_risk_quadrants <- spp_risk_quadrants %>%
  left_join(spp_risk_quadrants %>% 
              filter(quad == 'q1') %>%
              dplyr::select(popn_category, pct_q1 = pct_quad),
            by = 'popn_category') %>%
  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
         quad = str_replace(quad, 'q3', 'q3 area-matched'),
         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
  left_join(spp_list %>%
              dplyr::select(popn_category, cat_rank = category_score) %>%
              unique() %>%
              mutate(cat_rank = ifelse(popn_category == 'DD', -1, cat_rank)),
            by = 'popn_category')

### Plot the bar chart
barchart_spp_risk_quads <- ggplot(spp_risk_quadrants %>% 
                                  transform(popn_category = reorder(popn_category, cat_rank)), 
                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'count', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, 
            #size = 2, 
            color = 'grey20') +
  coord_flip() +
  labs(x = 'Risk category', 
       y = 'Relative proportion of species per quadrant', 
#       title = 'Risk category by quadrant',
       fill = 'Alignment')

@@

iucn_marine_file <- file.path(dir_data, 'spp_iucn_marine.csv')
spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)

message('Getting IUCN maps list')
iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp

@@

``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE}

@@

taxa_am_all_assessed <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(sciname %in% spp_iucn_marine$sciname) %>%
  mutate(class = str_to_title(class),
         n_am_tot = n()) %>%        ### length of whole AM maps list
  group_by(class) %>%
  summarize(n_am_class_ass  = n(),
            p_am_class_ass  = round((n_am_class_ass/first(n_am_tot) * 100), 2))

taxa_am_solo_assessed <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(sciname %in% spp_iucn_marine$sciname) %>%
  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
  mutate(class = str_to_title(class),
         n_am_solo = n()) %>%       ### length of AM only list
  group_by(class) %>%
  summarize(n_am_class_solo_ass = n(),
            p_am_class_solo_ass = round((n_am_class_solo_ass/first(n_am_solo) * 100), 2))

@@

``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE}
### bar charts comparing number of species in each class for each data set

>>>>

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           progress = TRUE)

@@

### define representative species from each quadrant, for over-plotting
# spp_quads <- c('q1' = 'Kajikia albida', 
#                'q2' = 'Galaxea cryptoramosa', 
#                'q4' = 'Acanthurus nigroris', 
#                'q4a' = 'Metanephrops australiensis', 
#                'q3' = 'Conus magnificus')
# 
### truncate the species names for the representative species
# spp_list_label <- spp_list %>%
#   filter(sciname %in% spp_rando) %>%
#   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
#   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
#   dplyr::select(-g, -s)

@@

spp_list <- spp_list %>% 
  mutate(data_poor = ifelse(occurcells < 10, '< 10', '>= 10'))
### mongo plot time
scatter_arearatio_vs_overlap2 <- ggplot(spp_list %>% 
                                         transform(popn_category = reorder(popn_category, category_score)),
                                       aes(x = area_ratio, y = sm_perc, 
                                           color = data_poor,
                                           size = lg_area_pct)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
  geom_point(alpha = .6)

### Manage scales for color and size 
scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 +
  scale_colour_manual(values = c('>= 10' = 'green4', 
                                 '< 10'  = 'orangered2')) + 
  scale_size_continuous(limits = c(0, 100),
                        breaks = c( 0,    25,    50,    75,    100), 
                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
  scale_x_continuous(expand = c(0, 0), limits = c(0, 100)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))

### here are median labels:
scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 +
  annotate(geom = 'text',
           x = area_ratio_quartiles[2], y = 5,
           hjust = 0, vjust = 0,
           color = 'grey30', 
           size = 2,
           fontface = 'bold.italic', angle = 90,
           label = sprintf('Median = %s%%', round(area_ratio_quartiles[2], 1))) +
  annotate(geom = 'text',
           x = 5, y = pct_overlap_quartiles[2],
           hjust = 0, vjust = 0,
           color = 'grey30', 
           size = 2,
           fontface = 'bold.italic', angle = 0,
           label = sprintf('Median = %s%%', round(pct_overlap_quartiles[2], 1)))

scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 +
  labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
#       title = 'Distribution alignment vs Extent alignment',
       color = 'AM model cells', size = '% ocean range')

print(scatter_arearatio_vs_overlap2)

@@

spp_list_quads <- spp_list1 %>%
  mutate(quad = NA,
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q1$sciname, 'q1', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q2$sciname, 'q2', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q3$sciname, 'q3', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
write_csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'))

@@

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1)) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey20') +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Relative proportion of species per quadrant', 
#       title = 'Taxonomic group by quadrant',
       fill = 'Alignment')

@@

write_csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'))

@@

### Summarize quadrant species list by species group (e.g. CORALS1)
spp_risk_q1 <- spp_list_q1 %>%
  group_by(popn_category) %>%
  summarize(n_spp_q1 = n())
spp_risk_q2 <- spp_list_q2 %>%
  group_by(popn_category) %>%
  summarize(n_spp_q2 = n())
spp_risk_q3 <- spp_list_q3 %>%
  group_by(popn_category) %>%
  summarize(n_spp_q3 = n())
spp_risk_q4 <- spp_list_q4 %>%
  group_by(popn_category) %>%
  summarize(n_spp_q4 = n())

### Join species group by quadrant into a single dataframe
spp_risk_quadrants <- spp_risk_q1 %>%
  full_join(spp_risk_q2, by = c('popn_category')) %>%
  full_join(spp_risk_q3, by = c('popn_category')) %>%
  full_join(spp_risk_q4, by = c('popn_category')) %>%
  gather(quad, n_spp_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
  group_by(popn_category) %>%
  mutate(n_tot_cat = sum(n_spp_quad)) %>%
  ungroup() %>%
  mutate(quad = str_replace(quad, 'n_spp_', ''),
         pct_quad   = n_spp_quad/n_tot_cat,
         pct_quad   = ifelse(is.na(pct_quad),   0, pct_quad))

spp_risk_quadrants1 <- spp_risk_quadrants %>%
  left_join(spp_risk_quadrants %>% 
              filter(quad == 'q1') %>%
              dplyr::select(popn_category, pct_q1 = pct_quad),
            by = 'popn_category') %>%
  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
         quad = str_replace(quad, 'q3', 'q3 area-matched'),
         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
  left_join(spp_list %>%
              dplyr::select(popn_category, cat_rank = category_score) %>%
              unique() %>%
              mutate(cat_rank = ifelse(popn_category %in% c('DD', 'NE'), -1, cat_rank)),
            by = 'popn_category')

### Plot the bar chart
barchart_spp_risk_quads <- ggplot(spp_risk_quadrants1 %>% 
                                  transform(popn_category = reorder(popn_category, cat_rank)), 
                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1)) +
  geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey20') +
  coord_flip() +
  labs(x = 'Risk category', 
       y = 'Relative proportion of species per quadrant', 
#       title = 'Risk category by quadrant',
       fill = 'Alignment')

@@

iucn_file <- file.path(dir_data, 'spp_iucn.csv')
spp_iucn <- read.csv(iucn_file, stringsAsFactors = FALSE)

message('Getting IUCN maps list')
iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
  filter(sciname %in% spp_iucn$sciname)   ### ditch freshwater spp

@@

``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}

@@

taxa_am_all_assessed <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(sciname %in% spp_iucn$sciname) %>%
  mutate(class = str_to_title(class),
         n_am_tot = n()) %>%        ### length of whole AM maps list
  group_by(class) %>%
  summarize(n_am_class_ass  = n(),
            p_am_class_ass  = round((n_am_class_ass/first(n_am_tot) * 100), 2))

taxa_am_solo_assessed <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(sciname %in% spp_iucn$sciname) %>%
  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
  mutate(class = str_to_title(class),
         n_am_solo = n()) %>%       ### length of AM only list
  group_by(class) %>%
  summarize(n_am_class_solo_ass = n(),
            p_am_class_solo_ass = round((n_am_class_solo_ass/first(n_am_solo) * 100), 2))

@@

``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
### bar charts comparing number of species in each class for each data set

### 2016-02-12 data_explore/data_explore_am_thresh.Rmd
updated plots with newly cleaned data

thresh_df_summary <- threshold_df %>%
  group_by(thresh) %>%
  summarize(mean_prop   = mean(proportion, na.rm = TRUE),
            median_prop = median(proportion, na.rm = TRUE),
            quant09 = quantile(proportion, .09),
            quant25 = quantile(proportion, .25),
            quant75 = quantile(proportion, .75),
            quant91 = quantile(proportion, .91))

>>>>

thresh_df_summary <- threshold_df %>%
  group_by(thresh) %>%
  summarize(mean_prop   = mean(proportion, na.rm = TRUE),
            median_prop = median(proportion, na.rm = TRUE),
            quant09 = quantile(proportion, .05),
            quant25 = quantile(proportion, .25),
            quant75 = quantile(proportion, .75),
            quant91 = quantile(proportion, .95))

### 2016-02-12 data_explore/data_prep.Rmd
oops, write the processed cell files
updated plots with newly cleaned data

gather spp_all, spp_am, spp_am_v_iucn, spp_list_w_area_all and _trimmed into data folder.

@@

### filter out spatial_source == 'am' that also have parent_sid (i.e.
### these are AM-mapped subpopulations); then: 
### select away the iucn_subpop, parent_sid, and subpop_sid (among others),
### these are creating false unique rows.
spp_all <- spp_all_remote_clean %>%
  filter(!(spatial_source == 'am' & !is.na(parent_sid))) %>%
  dplyr::select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
         popn_category, spp_group, id_no, spatial_source, name_verified) %>%
  unique()
### At this point, still some duped scinames - these are ones with multiple IUCN
### SIDs that are not designated as parents or subpops.  We will brute-force
### this by just eliminating the duplicated rows, keeping only the first instance.
### By arranging by iucn_sid, we will keep the lowest value, which seems to
### generally correspond to the parent...
spp_all <- spp_all %>%
  arrange(iucn_sid)
spp_all <- spp_all[!duplicated(spp_all$sciname), ]

@@

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
message('Building IUCN species to cell table.  This might take a few minutes.')
iucn_map_files      <- file.path(dir_anx, 'iucn_intersections', list.files(file.path(dir_anx, 'iucn_intersections')))
iucn_spp_cells <- lapply(iucn_map_files, data.table::fread) %>%
  bind_rows()
iucn_spp_cells <- iucn_spp_cells %>%
  rename(iucn_sid = id_no, loiczid = LOICZID) %>%
  mutate(iucn_sid = as.integer(iucn_sid),
         presence = as.integer(presence),
         loiczid  = as.integer(loiczid)) %>%
  filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
  filter(presence != 5)
### length(unique(iucn_spp_cells$iucn_sid)) results in 2013 species found...
### some species with no cells reported.  2249 listed in the am v iucn file.

message('Loading AquaMaps species to cell table.  This might take a few minutes.')
am_spp_cells <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
                         showProgress = TRUE)
am_spp_cells <- am_spp_cells %>%
  rename(am_sid = speciesid, prob = probability) %>%
  filter(am_sid %in% spp_am_v_iucn$am_sid)
### length(unique(am_spp_cells$am_sid)) results in 2283 species found - great!

@@

spp_list <- spp_list %>%
  mutate(sm_perc = ifelse(!is.na(sm_perc), 0, sm_perc))
  # filter(!is.na(sm_perc)) %>%
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?
spp_list <- spp_list %>%
  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
                     stringsAsFactors = FALSE),
            by = 'spp_group') %>%
  left_join(read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
                   stringsAsFactors = FALSE) %>% 
              select(am_sid = speciesid, occurcells),
            by = 'am_sid')

>>>>

gather spp_all, spp_am, spp_iucn, spp_iucn_maps, and spp_am_raw into local data folder.

@@

### filter out spatial_source == 'am' that also have parent_sid (i.e.
### these are AM-mapped subpopulations); then: 
### select away the iucn_subpop, parent_sid, and subpop_sid (among others),
### these are creating false unique rows.
spp_all <- spp_all_remote_clean %>%
  filter(!(spatial_source == 'am' & !is.na(parent_sid))) %>%
  dplyr::select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
                popn_category, category_score, spp_group, 
                id_no, spatial_source, name_verified) %>%
  unique()
### At this point, still some duped scinames - these are ones with multiple IUCN
### SIDs that are not designated as parents or subpops.  We will brute-force
### this by just eliminating the duplicated rows, keeping only the first instance.
### By arranging by iucn_sid, we will keep the lowest value, which seems to
### generally correspond to the parent...
spp_all <- spp_all %>%
  arrange(iucn_sid)
spp_all <- spp_all[!duplicated(spp_all$sciname), ]

@@

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
message('Building IUCN species to cell table.  This might take a few minutes.')
iucn_map_files <- file.path(dir_anx, 'iucn_intersections', list.files(file.path(dir_anx, 'iucn_intersections')))
iucn_spp_cells <- lapply(iucn_map_files, data.table::fread) %>%
  bind_rows()
iucn_spp_cells <- iucn_spp_cells %>%
  rename(iucn_sid = id_no, loiczid = LOICZID) %>%
  mutate(iucn_sid = as.integer(iucn_sid),
         presence = as.integer(presence),
         loiczid  = as.integer(loiczid)) %>%
  filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
  filter(presence != 5)
### length(unique(iucn_spp_cells$iucn_sid)) results in 2013 species found...
### some species with no cells reported.  2249 listed in the am v iucn file.
write_csv(iucn_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'))

message('Loading AquaMaps species to cell table.  This might take a few minutes.')
am_spp_cells <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
                         showProgress = TRUE)
am_spp_cells <- am_spp_cells %>%
  rename(am_sid = speciesid, prob = probability) %>%
  filter(am_sid %in% spp_am_v_iucn$am_sid)
### length(unique(am_spp_cells$am_sid)) results in 2283 species found - great!
write_csv(am_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))

@@

spp_list <- spp_list %>%
  mutate(sm_perc = ifelse(is.na(sm_perc), 0, sm_perc)) %>%
  filter(!area_iucn == 0)
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?

### attach group name text field, and occurcells field
spp_list <- spp_list %>%
  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
                     stringsAsFactors = FALSE),
            by = 'spp_group') %>%
  left_join(read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
                   stringsAsFactors = FALSE) %>% 
              select(am_sid = speciesid, occurcells),
            by = 'am_sid')

cat_score_df <- data.frame(popn_category  = c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE'),
                           category_score = c( 0.0,  0.2,  0.4,  0.6,  0.8,  1.0,   NA,  NA))
spp_list <- spp_list %>%
  left_join(cat_score_df, by = 'popn_category')
